---
title: "Peer Assessment II"
output: html_document
---

# Background

As a statistical consultant working for a real estate investment firm, your task is to develop a model to predict the selling price of a given home in Ames, Iowa. Your employer hopes to use this information to help assess whether the asking price of a house is higher or lower than the true value of the house. If the home is undervalued, it may be a good investment for the firm.

# Training Data and relevant packages

In order to better assess the quality of the model you will produce, the data have been randomly divided into three separate pieces: a training data set, a testing data set, and a validation data set. For now we will load the training data set, the others will be loaded and used later.

```{r load, message = FALSE}
load("ames_train.Rdata")
```

Use the code block below to load any necessary packages

```{r packages, message = FALSE}
library(MASS)
library(statsr)
library(tidyverse)
library(BAS)
library(reshape2)
library(ggpubr)
library(ModelMetrics)
```

## Part 1 - Exploratory Data Analysis (EDA)

When you first get your data, it's very tempting to immediately begin fitting models and assessing how they perform.  However, before you begin modeling, it's absolutely essential to explore the structure of the data and the relationships between the variables in the data set.

Do a detailed EDA of the ames_train data set, to learn about the structure of the data and the relationships between the variables in the data set (refer to Introduction to Probability and Data, Week 2, for a reminder about EDA if needed). Your EDA should involve creating and reviewing many plots/graphs and considering the patterns and relationships you see. 

After you have explored completely, submit the three graphs/plots that you found most informative during your EDA process, and briefly explain what you learned from each (why you found each informative).

* * *

```{r creategraphs}
melted <- ames_train %>% 
        select(price, Lot.Area, Bedroom.AbvGr,Overall.Qual,Year.Built) %>%
        melt(id.vars=1)

g<-ggplot(melted, aes(x = value, y = price)) + theme_bw()
g<- g + facet_wrap(~variable,scales = 'free_x') + geom_point() 
g<- g + geom_smooth(formula=y~x,method = 'lm')
g
```

From this plot we can see how some of the potential explanatory variables appears to be better as single predictors than others.

```{r}
trns <- ames_train %>% 
        select(price,area) %>% 
        mutate(lnprice = log(price),lnarea = log(area))

melted2 <- trns %>% melt(id.vars = c(1,3)) 

g0<- ggplot(melted2,aes(x = value)) + theme_bw() + facet_wrap(~variable,scales = 'free_x')
g1<- g0 + geom_point(aes(y = price)) + geom_smooth(aes(y=price),formula=y~x,method = 'lm')
g2<- g0 + geom_point(aes(y = lnprice)) + geom_smooth(aes(y=lnprice),formula=y~x,method = 'lm')

ggarrange(g1,g2,ncol = 1)
```

This plot shows how some transformations can significantly help the modeling process, making the relationship more linear between the variables.

```{r}
ggplot(ames_train, aes(x = Neighborhood, y = price)) + geom_boxplot(aes(fill = Neighborhood)) + 
        theme_bw() + 
        scale_x_discrete(label = NULL)
```

This is a important graph that shows how price can vary between neighborhoods. We can identify some neighborhoods where the variations are particularly big.

* * *

## Part 2 - Development and assessment of an initial model, following a semi-guided process of analysis

### Section 2.1 An Initial Model
In building a model, it is often useful to start by creating a simple, intuitive initial model based on the results of the exploratory data analysis. (Note: The goal at this stage is **not** to identify the "best" possible model but rather to choose a reasonable and understandable starting point. Later you will expand and revise this model to create your final model.

Based on your EDA, select *at most* 10 predictor variables from “ames_train” and create a linear model for `price` (or a transformed version of price) using those variables. Provide the *R code* and the *summary output table* for your model, a *brief justification* for the variables you have chosen, and a *brief discussion* of the model results in context (focused on the variables that appear to be important predictors and how they relate to sales price).

* * *

```{r fit_model}
ames_train %>% mutate(Total.Bsmt.SF = ifelse(is.na(Total.Bsmt.SF),0,Total.Bsmt.SF), Garage.Cars = ifelse(is.na(Garage.Cars),0,Garage.Cars)) -> ames_train

model.full <- lm(log(price) ~ Overall.Qual +   
                  log(Total.Bsmt.SF + 1) + Garage.Cars + log(area) + 
                  Half.Bath + Bedroom.AbvGr + Year.Built +
                  log(Lot.Area) +  Central.Air + Overall.Cond,
                 data = ames_train)
summary(model.full)

```

Those are the variables that showed highest correlations at EDA. I choose the highest allowed number of them to start with because I can further do a model selection from here, and in this case, it is good to have a great quantity of models to test. 

From the estimated coefficients we can see Log(area) and Log(Lot.Area) appear to be the ones with highest influence per unit increase, with Log(area) being by far the highest.

* * *

### Section 2.2 Model Selection

Now either using `BAS` another stepwise selection procedure choose the "best" model you can, using your initial model as your starting point. Try at least two different model selection methods and compare their results. Do they both arrive at the same model or do they disagree? What do you think this means?

* * *

```{r model_select}
model.bas <- bas.lm(log(price) ~ Overall.Qual +   
                  log(Total.Bsmt.SF + 1) + Garage.Cars + log(area) + 
                  Half.Bath + Bedroom.AbvGr + Year.Built +
                  log(Lot.Area) +  Central.Air + Overall.Cond,
                 data = ames_train)

summary(model.bas)

model.AIC <- stepAIC(model.full, k = 2, trace = F)
summary(model.AIC)
```

The best model, either using `BAS` or `stepAIC`, was the same. All the selected variables remained. This shows how strong is the relationship between those exploratory variable and the response variable, making them important predictors.


* * *

### Section 2.3 Initial Model Residuals
One way to assess the performance of a model is to examine the model's residuals. In the space below, create a residual plot for your preferred model from above and use it to assess whether your model appears to fit the data well. Comment on any interesting structure in the residual plot (trend, outliers, etc.) and briefly discuss potential implications it may have for your model and inference / prediction you might produce.

* * *

```{r model_resid}
plot(model.full, which = 1, add.smooth = F,
ask = F, pch = 16, sub.caption="", caption="")
abline(a = 0, b = 0, col = "darkgrey", lwd = 2)
```

The plot shows that overall, we have constant variance of residuals, except for the presence of some potential outliers, that being observations 428, 181 and 310. 

This is an assumption that needs to be checked for the reliability of a linear model, which we assume to have residuals following a normal distribution with mean zero and variance $\sigma^2$.

We can see bellow those potential outliers actually have 0 probability of being outliers, so everything is okay.

```{r}
outliers <- Bayes.outlier(model.full,k=2)

outliers$prob.outlier[c(428,181,310)] %>% round(3)
```




* * *

### Section 2.4 Initial Model RMSE

You can calculate it directly based on the model output. Be specific about the units of your RMSE (depending on whether you transformed your response variable). The value you report will be more meaningful if it is in the original units (dollars).

* * *

```{r model_rmse}
pred.train <- exp(model.full$fitted.values)

rmse(ames_train$price,pred.train)

```

* * *

### Section 2.5 Overfitting 

The process of building a model generally involves starting with an initial model (as you have done above), identifying its shortcomings, and adapting the model accordingly. This process may be repeated several times until the model fits the data reasonably well. However, the model may do well on training data but perform poorly out-of-sample (meaning, on a dataset other than the original training data) because the model is overly-tuned to specifically fit the training data. This is called “overfitting.” To determine whether overfitting is occurring on a model, compare the performance of a model on both in-sample and out-of-sample data sets. To look at performance of your initial model on out-of-sample data, you will use the data set `ames_test`.

```{r loadtest, message = FALSE}
load("ames_test.Rdata")
```

Use your model from above to generate predictions for the housing prices in the test data set.  Are the predictions significantly more accurate (compared to the actual sales prices) for the training data than the test data?  Why or why not? Briefly explain how you determined that (what steps or processes did you use)?

* * *

```{r initmodel_test}
pred.test <- exp(predict(model.full,ames_test))

rmse(ames_test$price,pred.test)
```

We can see the rmse for out of sample data is lower than the in sample, so the model generalizes well. It's good to note here that the high in sample rmse is probably due to some extreme value present at the in sample but not present at the out of sample data. This should be analyzed further.

* * *

**Note to the learner:** If in real-life practice this out-of-sample analysis shows evidence that the training data fits your model a lot better than the test data, it is probably a good idea to go back and revise the model (usually by simplifying the model) to reduce this overfitting. For simplicity, we do not ask you to do this on the assignment, however.

## Part 3 Development of a Final Model

Now that you have developed an initial model to use as a baseline, create a final model with *at most* 20 variables to predict housing prices in Ames, IA, selecting from the full array of variables in the dataset and using any of the tools that we introduced in this specialization.  

Carefully document the process that you used to come up with your final model, so that you can answer the questions 1

### Section 3.1 Final Model

Provide the summary table for your model.

* * *

```{r model_playground}

model.final <- lm(log(price) ~ Overall.Qual +   
                  log(Total.Bsmt.SF + 1) + Garage.Cars + log(area) + 
                  Bedroom.AbvGr + Year.Built + 
                  log(X2nd.Flr.SF + 1) +
                  log(Lot.Area) +  Central.Air + Overall.Cond,
                 data = ames_train)

summary(model.final)
```

* * *

### Section 3.2 Transformation

Did you decide to transform any variables?  Why or why not? Explain in a few sentences.

* * *

It was necessary to do some log transformation of some variables related with size due to skewness. Those transformations helps the modeling process because it makes the variables more normally distributed. 

```{r model_assess}
```

* * *

### Section 3.3 Variable Interaction

Did you decide to include any variable interactions? Why or why not? Explain in a few sentences.

* * *

It wasn't necessary. We already have a good set of variables to work with.

```{r model_inter}
```

* * *

### Section 3.4 Variable Selection

What method did you use to select the variables you included? Why did you select the method you used? Explain in a few sentences.

* * *

It was used Bayesian Adaptive Sampling, `BAS`, and the Highest Probability Model was selected. The Bayesian approach is a very reliable method of variable selection, it calculates the Marginal Posterior Inclusion Probabilities of each variable, using Bayesian Statistics.

```{r model_select_final}
model.final.bas<-bas.lm(log(price) ~ Overall.Qual + log(Garage.Area + 1) +   
                                log(Total.Bsmt.SF + 1) + Garage.Cars +
                                log(area) + Full.Bath + Half.Bath + 
                                Bedroom.AbvGr + Year.Built + 
                                log(X1st.Flr.SF+1) +
                                log(X2nd.Flr.SF + 1) +
                                log(Lot.Area) +  Central.Air + Overall.Cond,
                        data = ames_train)
summary(model.final.bas)
```

* * *

### Section 3.5 Model Testing

How did testing the model on out-of-sample data affect whether or how you changed your model? Explain in a few sentences.

* * *

To avoid overfitting problems, every model was evaluated on out-of-sample. This way we guarantee that we would choose the model that better generalizes.

* * *

## Part 4 Final Model Assessment

### Section 4.1 Final Model Residual

For your final model, create and briefly interpret an informative plot of the residuals.

* * *

The plot bellow shows that overall, we have constant variance of residuals, except for the presence of some potential outliers, that being observations 428, 181 and 310. 

This is an assumption that needs to be checked for the reliability of a linear model, which we assume to have residuals following a normal distribution with mean zero and variance $\sigma^2$.

```{r}
plot(model.final.bas, which = 1, add.smooth = F,
ask = F, pch = 16, sub.caption="", caption="")
abline(a = 0, b = 0, col = "darkgrey", lwd = 2)
```

We can see bellow those potential outliers actually have 0 probability of being outliers, so everything is okay.

```{r}
outliers <- Bayes.outlier(model.final,k=2)

outliers$prob.outlier[c(428,181,310)] %>% round(3)
```


* * *

### Section 4.2 Final Model RMSE

For your final model, calculate and briefly comment on the RMSE.

* * *

As can be seen bellow, rmse of the final model is bellow the one for the initial model. So we were able to lower the out of sample error with the new variables selected, making the model able to generalize more.

```{r model_testing}

pred.final.test <- exp(predict(model.final.bas,ames_test,estimator = 'HPM')$fit)

rmse(ames_test$price,pred.test)
rmse(ames_test$price,pred.final.test)

```

* * *

### Section 4.3 Final Model Evaluation

What are some strengths and weaknesses of your model?

* * *

The model seems to generalize well enough.

* * *

### Section 4.4 Final Model Validation

Testing your final model on a separate, validation data set is a great way to determine how your model will perform in real-life practice. 

You will use the “ames_validation” dataset to do some additional assessment of your final model. Discuss your findings, be sure to mention:
* What is the RMSE of your final model when applied to the validation data?  
* How does this value compare to that of the training data and/or testing data?
* What percentage of the 95% predictive confidence (or credible) intervals contain the true price of the house in the validation data set?  
* From this result, does your final model properly reflect uncertainty?

```{r loadvalidation, message = FALSE}
load("ames_validation.Rdata")
```

* * *

```{r model_validate}
pred.final.val <- exp(predict(model.final.bas,
                              ames_validation,
                              estimator = 'HPM')$fit)

rmse(ames_train$price,exp(model.final$fitted.values))
rmse(ames_test$price,pred.final.test)
rmse(ames_validation$price,pred.final.val)

inter.final.val <- predict(model.final.bas, 
                           ames_validation, 
                           estimator="HPM",
                           prediction=TRUE, 
                           se.fit=TRUE)

out <- as.data.frame(cbind(exp(confint(inter.final.val)),
                           price = ames_validation$price))

colnames(out)[1:2] <- c("lwr", "upr")
pred.final.val.coverage <- out %>% summarize(cover = sum(price >= lwr & price <= upr)/n())
pred.final.val.coverage*100 %>% round(3)

```

We can see from those results the model generalizes well and properly reflect uncertainty, with 97.64% of the validation data covered.

* * *

## Part 5 Conclusion

Provide a brief summary of your results, and a brief discussion of what you have learned about the data and your model. 

* * *

We were able to build a very reliable model on out-of-sample data that properly reflects uncertainty, using `BAS` and selecting the model with highest posterior probability. 

Some variables, including the response variable itself, needed to go through some log transformation, due to the presence of skewness.


* * *
